Child Signals of Need in Utila, Honduras

### 2024-08-31 Paper.qmd
Analysis script: moved some recodings to data package, updated and fixed plots and updated alloparenting (dependent variable) model.
Paper: Added to results and methods. Also updated and added some citations.

---
title: "Child Signals of Need in Utila, Honduras"
execute: 
  echo: false
  warning: false
format: html
editor: visual
tbl-cap-location: bottom
bibliography: signaling.bib
---

@@

Despite the complexity of childhood and adolescence and their importance to adult outcomes, the behavioral strategies of children are often overlooked, especially when it comes to how they elicit support from parents and other caretakers [@gottlieb2000; @wells2003; @hirschfeld2002]. To gain support, children use signals to communicate their desires to parents. Infant signals of need, particularly infant cries, have been extensively studied in a range of populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], yet the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages. In particular, if and when children signal need with costly signals [@grafen1990; @spence1973; @zahavi1975a] has rarely been investigated (for one exception, see [@syme2023]).

@@

Conflicts of interest between children and parents may also involve how much children invest in the household. As previously mentioned, each individual child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974]. This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food. Such a dynamic has been found among the Chuukese, with both male and female children reporting conflict with their families over children’s desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003]. Similarly, Syme and Hagen [@syme2023] found greater numbers of siblings were associated with increased likelihoods of severe conflict in a retrospective study with Chuukese-Americans. This may also be consistent with parents pressuring older children to invest in the household; however, there are other ways in which limited resources can result in behavioral conflict [@syme2023].

@@

Withdrawal has been suggested to be an underappreciated way those with low social power may credibly signal need [@syme2023]. Due to their lower status, children have limited ways to change how others treat them. However, as children 1) are their parents’ path to direct fitness and 2) can help family members through allocare and other forms of labor, they provide benefits that are valuable and hard to replace. By withholding these benefits or putting the future of these benefits into question, a child simultaneously delays the benefits they get from this cooperation while imposing costs on those with a fitness interest in the child. As these costs are likely lower for those truly in need than those who are not, withdrawal may be a credible way to signal need if the costs to faking the signal make it unprofitable. In these cases, the end result is analogous to a labor strike, in which the signaler conveys to others their situation demands more support, and if they want to continue receiving benefits from the child, they need to upregulate their support of the child.

@@

|                                 |                                                                                                                   |
|-------------------------|----------------------------------------------|
| Variable                        | Survey question wording                                                                                           |
| Crying Frequency                | How often does your child cry?                                                                                    |
| Sad Frequency                   | How often is your child sad?                                                                                      |
| Tantrum Frequency               | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
| Running Away Frequency          | How often does your child leave the house without telling you when they get mad?                                  |
| Parent/Child Conflict Frequency | How often do you have conflicts with your child?                                                                  |
| Child Alloparenting Frequency   | How often does your child take care of their other siblings?                                                      |

@@

From these count variables, we created three additional signaling variables to be used, primarily, as our main dependent variables in analyses. The frequency of signaling was the sum of the frequencies of child sadness, crying, and temper tantrums. Maximum signal frequency was the frequency of the most common of these three signals. Signal cost was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., Signal cost = frequency of sadness + frequency of crying * 2 + frequency of temper tantrums * 3). Although we collected data on the frequency of running away, it was not included in the creation of any of these variables due to uncertainty over whether it was likely to be a credible signal due to the relative safety of Utila compared to other areas in which running away has been found to be a successful strategy for children.

@@

|                     |                                                                                                  |
|-------------------------|----------------------------------------------|
| Variable            | Survey question wording                                                                          |
| Relative Need       | How much time, effort, and resources does this child need compared to your other children?       |
| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |

@@

We also asked caretakers about the last time each participating child was sad or cried. This involved asking what the caretaker thought caused the emotion or display and recording their answer in a text box. We then asked how the parent responded to their child. The researcher giving the survey would then bin responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective). For the full list of options see Table S1.

From this data, we created a parental response variable, which was coded as 1 for a positve response, 0 for a neutral response, and -1 for a negative response (all from the child's perspective). MG also coded each cause for the presence or absence of 17 various elements. Of these 4 were use in analyses: whether or not the cause involved 1) the child being asked to do an unwanted task, 2) punishment, 3) conflict within the family, and 4) discomfort, pain, injury, or illness.

### Child anthropometrics

We computed each child's height, weight, and BMI Zscores with respect to the CDC child growth standards (cite).

@@

|                |                                                                                       |
|-------------------------|----------------------------------------------|
| Variable       | Survey question wording                                                               |
| Currenthealth1 | If an illness is 'going around', my child will get it.                                |
| Currenthealth2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| Currenthealth3 | My child’s immune system protects him/her from most illnesses that other kids get.    |

@@

The survey also included many demographic questions about the survey taker and their households. Those used for this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide allocare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income. This data allowed us to calculate (1) the number of children within the household and (2) the number of adults within the household.

OLD AND CURRENTLY NOT USING

We also calculated the number of individuals who engaged in childcare within the household, by combining the number of adults who provide care within the household with data from child alloparenting frequencies. Due to the fact the adult childcare variable was a count of helpers and the child alloparenting variable was frequency based, we created 3 total alloparenting variables as a robustness check. "OtherAlloparents" is the sum of the number of adults who provide childcare plus the sum of other children who engaged in alloprarenting at least "more than once a week but not daily." The second and third involved summing the alloparenting frequency (AlloparentingFreqN) of other children within the household. We then summed this with the frequency of adult childcare, which either scored each adult who was listed as providing childcare as helping 30 times per month (AlloparentingFreqNa) or 60 times per month (AlloparentingFreqNb).

@@

We fit generalized linear mixed models for exploratory analyses with signal frequencies and summed signal cost as dependent variables. As our signaling variables invovled over-dispersed count data, we used negative binomial regression. This was done in R with the GLMM package [@brooks2017], with householdID as a random effect.

For the models examining 1) parental responses to the last reported time each participating child was sad or cried, 2) the relative neediness of the child, and 3) the investment a child received compared to the other children in the household, we fit ordinal logistic regression models using two methods. The first used the polr function in the MASS package [@venables2002]. The results were then interpreted with clustered standard errors for householdID. Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS with householdID as a random effect.

@@

```{r sumtab_children}
#| label: tbl-sumtab_children
#| tbl-cap: Summary statistics for child study variables.
vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  Currenthealth1 = "Currenthealth1",
  Currenthealth2 = "Currenthealth2",
  Currenthealth3 = "Currenthealth3",
  CurrentHealthMean = "Mean of current health measures"
)

@@

# CONFIRM: should this be household level?
```{r sumtab_adults}
#| label: tbl-sumtab_adults
#| tbl-cap: Summary statistics for adult and household study variables.

vars_c <- c(
  CaregiverAge = "Age (years)",
  EducationLevelYears = "Education level (years)",
  IncomeCategoryN = "Household income (Lempira)"
)

@@

Child sadness, crying, and temper tantrums were reasonably common. Caregivers reported that (`r sum(na.omit(modeldf$SadFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\]) of children were sad at least once a week, (`r sum(na.omit(modeldf$CryFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\]) cried at least more than once a week, and (`r sum(na.omit(modeldf$TantrumFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]) threw temper tantrums at least once a week. A small subset of children engaged in these signals/states at least daily (sadness = `r sum(na.omit(modeldf$SadFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$SadFreqN >= 30)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$CryFreqN >= 30)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 30)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]). However, many children were reported to never engage in these signals/states (sadness = `r sum(na.omit(modeldf$SadFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$SadFreqN == 00)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$CryFreqN == 0)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN == 0)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\])

Conflict between children and caretakers was also common, with `r sum(na.omit(modeldf$ConflictFreqN >= 8))` children (`r signif(sum((na.omit(modeldf$ConflictFreqN >= 8)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%) having conflicts with parents more than once a week. As with the frequencies of various types of signaling, many children (`r sum(na.omit(modeldf$ConflictFreqN == 0))` \[`r signif(sum((na.omit(modeldf$ConflictFreqN == 0)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%\]) were described as never having conflicts with parents (stat).

The frequency of alloparenting effort was bimodally distributed, with `r sum(na.omit(modeldf$AlloparentingFreqN == 0))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN == 0)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `r sum(na.omit(modeldf$AlloparentingFreqN >= 30))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN >= 30)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to help with childcare at least once a day.

@@

Stacked bar plot showing the frequencies of child sadness, crying, and temper tantrums per month as reported by primary caretaker. Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day." 
:::

@@

Stacked bar plot showing the frequency of conflict between child and primary caregiver as reported by primary caretaker. Color corresponds to sex of the child. Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day." 
:::

@@

Stacked bar plot showing the frequency of conflict a child's alloparenting effort as reported by primary caretaker. Color corresponds to sex of the child. Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day."
:::

@@

# FIX: y axis extends farther for tantrum model (reducing this manually breaks error shading)

@@

@@

@@

The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-main_models for the corresponding regression table.
:::

For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child immune function. The slope of these relationships were curvilinear and positive for girls and flat or slightly negative for boys. In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys. See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.

>>>>

---
title: "Child Signals of Need in Utila, Honduras"
execute: 
  echo: false
  warning: false
format: html
lightbox: true
editor: visual
tbl-cap-location: bottom
bibliography: signaling.bib
---

@@

Despite the complexity of childhood and adolescence and their importance to adult outcomes, the behavioral strategies of children are often overlooked, especially when it comes to how they elicit support from parents and other caretakers [@gottlieb2000; @wells2003; @hirschfeld2002]. To gain support, children use signals to communicate their desires to parents. Infant signals of need, particularly infant cries, have been extensively studied in a range of populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], yet the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages. In particular, if and when children signal need with costly signals [@grafen1990; @spence1973; @zahavi1975a] has rarely been investigated [for one exception, see, @syme2023a].

@@

Conflicts of interest between children and parents may also involve how much children invest in the household. As previously mentioned, each individual child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974]. This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food. Such a dynamic has been found among the Chuukese, with both male and female children reporting conflict with their families over children’s desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003]. Similarly, Syme and Hagen [@syme2023a] found greater numbers of siblings were associated with increased likelihoods of severe conflict in a retrospective study with Chuukese-Americans. This may also be consistent with parents pressuring older children to invest in the household; however, there are other ways in which limited resources can result in behavioral conflict [@syme2023a].

@@

Withdrawal has been suggested to be an underappreciated way those with low social power may credibly signal need [@syme2023a]. Due to their lower status, children have limited ways to change how others treat them. However, as children 1) are their parents’ path to direct fitness and 2) can help family members through allocare and other forms of labor, they provide benefits that are valuable and hard to replace. By withholding these benefits or putting the future of these benefits into question, a child simultaneously delays the benefits they get from this cooperation while imposing costs on those with a fitness interest in the child. As these costs are likely lower for those truly in need than those who are not, withdrawal may be a credible way to signal need if the costs to faking the signal make it unprofitable. In these cases, the end result is analogous to a labor strike, in which the signaler conveys to others their situation demands more support, and if they want to continue receiving benefits from the child, they need to upregulate their support of the child.

@@

|                                 |                                                                                                                   |
|--------------------------|----------------------------------------------|
| Variable                        | Survey question wording                                                                                           |
| Crying Frequency                | How often does your child cry?                                                                                    |
| Sad Frequency                   | How often is your child sad?                                                                                      |
| Tantrum Frequency               | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
| Running Away Frequency          | How often does your child leave the house without telling you when they get mad?                                  |
| Parent/Child Conflict Frequency | How often do you have conflicts with your child?                                                                  |
| Child Alloparenting Frequency   | How often does your child take care of their other siblings?                                                      |

@@

From these count variables, we created three additional signaling variables to be used, primarily, as our main dependent variables in analyses. The frequency of signaling was the sum of the frequencies of child sadness, crying, and temper tantrums. Maximum signal frequency was the frequency of the most common of these three signals. Signal cost was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., Signal cost = frequency of sadness + frequency of crying \* 2 + frequency of temper tantrums \* 3). Although we collected data on the frequency of running away, it was not included in the creation of any of these variables due to uncertainty over whether it was likely to be a credible signal due to the relative safety of Utila compared to other areas in which running away has been found to be a successful strategy for children.

We also calculated the summed frequency of child alloparenting contributions by other children within the household.

@@

|                     |                                                                                                  |
|--------------------------|----------------------------------------------|
| Variable            | Survey question wording                                                                          |
| Relative Need       | How much time, effort, and resources does this child need compared to your other children?       |
| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |

@@

We asked caretakers about the last time each participating child was sad or cried. This involved asking what the caretaker thought caused the emotion or display and recording their answer in a text box. We then asked how the parent responded to their child. The researcher giving the survey would then bin responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective). For the full list of options see Table S1.

From this data, we created a parental response variable, which was coded as 1 for a positive response, 0 for a neutral response, and -1 for a negative response (all from the child's perspective). MG also coded each case for whether or not the cause had involved: 1) the child being asked to do an unwanted task, 2) punishment, 3) conflict within the family, and 4) discomfort, pain, injury, or illness.

@@

|                |                                                                                       |
|--------------------------|----------------------------------------------|
| Variable       | Survey question wording                                                               |
| Currenthealth1 | If an illness is 'going around', my child will get it.                                |
| Currenthealth2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| Currenthealth3 | My child’s immune system protects him/her from most illnesses that other kids get.    |

@@

The survey also included many demographic questions about the survey taker and their households. Those used for this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide allocare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income. This data allowed us to calculate the number of children (total, younger, and older) and adults within the household.

## Child anthropometrics

xxxxxxxxxxDescribe camps brieflyxxxxxxxxxxxxxx - include dates and how we handled those

Triceps and subscapular skinfold thicknesses were measured using skinfold calipers. Measurements were taken directly on the skin. The triceps measurement occurred at the midpoint of the upper arm. The subscapular measurement occured immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna].

We computed each child's height, weight, and BMI Zscores with respect to the CDC child growth standards [@kuczmarski2002].

@@

We fit generalized linear mixed models for exploratory analyses with signal frequencies and summed signal cost as dependent variables. As our signaling variables invovled over-dispersed count data, we used negative binomial regression. This was done in R with the GLMM package [@brooks2017], with household as a random effect. Control variables differed in some models, but always include child age and sex as controls. Specific control variables are mentioned in the results

For the models examining 1) parental responses to the last reported time each participating child was sad or cried, 2) the relative neediness of the child, and 3) the investment a child received compared to the other children in the household, we fit ordinal logistic regression models using two methods. The first used the polr function in the MASS package [@venables2002]. The results were then interpreted with clustered standard errors for household. Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS with household as a random effect.

@@

`r dplyr::n_distinct(modeldf$householdID)` primary caretakers took the survey and provided signaling data for at least one child. `r nrow(modeldf)` children had frequency data for at least 1 signal, with `r nrow(modeldf_FULLSIG)` having data for all 3. Among those with at least one record of signaling frequency `r sex_table[1]` were female and `r sex_table[2]` were male. Anthropometric and signaling data exist for `r nrow(modeldf_FULLANTHROPOMETRICS)` children, of whom `r sex_table2[1]` were female and `r sex_table2[2]` were male. For summary statistics of child-level variables see @tbl-sumtab_children. For summary statistics of caretaker and household-level variables see @tbl-sumtab_adults.

```{r sumtab_children}
#| label: tbl-sumtab_children
#| tbl-cap: Summary statistics for child study variables.
vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  Currenthealth1 = "Currenthealth1",
  Currenthealth2 = "Currenthealth2",
  Currenthealth3 = "Currenthealth3",
  CurrentHealthMean = "Mean of current health measures",
  HeightZ = "Height Z-score (CDC)",
  WeightZ = "Weight Z-score (CDC)",
  BMIZ = "BMI Z-score (CDC)",
  SubscapMean = "Subscapular skinfold thickness",
  SubscapR = "Subscapular skinfold residuals",
  TricepMean = "Tricep skinfold thickness",
  TricepR = "Tricep skinfold residuals",
  BodyFat =  "Body fat composite"
)

@@

```{r sumtab_adults}
#| label: tbl-sumtab_adults
#| tbl-cap: Summary statistics for adult and household study variables.

vars_c <- c(
  CaregiverAge = "Age (years)",
  EducationLevelYears = "Education level (years)",
  IncomeCategoryN = "Monthly household income (Lempira)",
  number_adults = "Adults in household",
  AdultsChildcare = "Adults who provide childcare"
)

kable(custom.summarize(caregivers, vars_c)) %>% kable_styling()

@@

Child sadness, crying, and temper tantrums were reasonably common. Caregivers reported that (`r sum(na.omit(modeldf$SadFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\]) children were sad at least once a week, (`r sum(na.omit(modeldf$CryFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\]) cried at least more than once a week, and (`r sum(na.omit(modeldf$TantrumFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]) threw temper tantrums at least once a week. A small subset of children engaged in these signals/states at least daily (sadness = `r sum(na.omit(modeldf$SadFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$SadFreqN >= 30)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$CryFreqN >= 30)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 30)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]). However, many children were reported to never engage in these signals/states (sadness = `r sum(na.omit(modeldf$SadFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$SadFreqN == 00)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$CryFreqN == 0)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN == 0)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]). See @fig-barplot_SignalFreq for plot showing signaling frequencies and @fig-signal_alluvial_plot for plot showing the number of individuals who engaged in specific combinations of signal frequencies..

Conflict between children and caretakers was also common, with `r sum(na.omit(modeldf$ConflictFreqN >= 8))` children (`r signif(sum((na.omit(modeldf$ConflictFreqN >= 8)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%) having conflicts with parents more than once a week. As with the frequencies of various types of signaling, many children (`r sum(na.omit(modeldf$ConflictFreqN == 0))` \[`r signif(sum((na.omit(modeldf$ConflictFreqN == 0)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%\]) were described as never having conflicts with parents. See @fig-barplot_ConflictFreq for plot showing the frequencies of conflict between caretakers and children.

The frequency of alloparenting effort was bimodally distributed, with `r sum(na.omit(modeldf$AlloparentingFreqN == 0))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN == 0)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `r sum(na.omit(modeldf$AlloparentingFreqN >= 30))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN >= 30)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to help with childcare at least once a day. See @fig-barplot_AlloparentingFreq for plot showing the frequencies child alloparenting effort.

@@

Stacked bar plot showing the frequencies of child sadness, crying, and temper tantrums per month as reported by primary caretaker. Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
:::

@@

Stacked bar plot showing the frequency of conflict between child and primary caregiver as reported by primary caretaker. Color corresponds to sex of the child. Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
:::

@@

Stacked bar plot showing the frequency of conflict a child's alloparenting effort as reported by primary caretaker. Color corresponds to sex of the child. Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
:::

@@

@@

## Signaling behavior is negatively associated with caregiver education

There were statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`r stats$msadH$EducationLevelYears$str`) when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction. See @tbl-main_models for regression statistics @fig-effect_edu for effects plots.

::: {#fig-effect_edu}
```{r}
signaling_data_plot_edu
```

The effect of years of caregiver education on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). See @tbl-main_models for the corresponding regression table.
:::

## Predictors of conflict between children and caretakers

The frequency of conflict between children and caretakers is negatively associated with child age (`r stats$mconflict$ChildAge$str`) and the number of adults who provide childcare in a household (`r stats$mconflict$AdultsChildcare$str`) when controlling for the sex of the child, household demographics, the number of adults who provide and do not contribute to childcare, an alloparenting effort by sex interaction, and the summed frequency of alloparenting from all other children in the household. See @tbl-conflict_dep_var for the corresponding regression table and @fig-effect_conflict_dep_var for effects plots.

There was also a negative relationship between conflict frequency and the number of older children, however, this was not statistically significant (`r stats$mconflict$OlderKids$str`). Similarly, the model predicted conflict to be more common between caretakers and males, but this also failed to reach statistical significance (`r stats$mconflict$SexMale$str`).

::: {#fig-effect_conflict_dep_var}
```{r}
signaling_data_plot_conflict_dep_var
```

The effect of A) child age and B) the number of adults in the household that contribute to childrcare on the frequency of conflict between caretaker and child. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-conflict_dep_var for the corresponding regression table.
:::

```{r}
#| label: tbl-conflict_dep_var
#| tbl-cap: Regression table for the model examining the frequency of conflict between child and caretaker as the dependent variable.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  CurrentHealthMean = "Percieved immunity",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)",
  RelativeNeed3 = "Relative need (w/in household)",
  UnwantedTask = "Unwanted task",
  Punishment21 = "Punishment",
  Family21 = "Family conflict",
  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
  CaregiverResponse = "Caregiver Response",
  OlderKids = "Older children",
  YoungerKids = "Younger children",
  AdultsNoChildcare = "Adults that do not provide childcare",
  AdultsChildcare = "Adults that provide childcare",
  OtherChildAlloparentingFreqN = "Other household child alloparetning freq."
)
modelsummary(mconflict, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

@@

#FIX coords do not work for this

@@

The effect of A) child age and B) the number of adults that contribute to childrecare within the household on the frequency of conflict between child and caretaker. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Rug corresponds to actual values of child alloparenting frequency with a small amount of jitter to reveal overlapped values. See @tbl-conflict_dep_var for the corresponding regression table.
:::

For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child immune function. The slope of these relationships were positive for girls and flat or slightly negative for boys. In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys. See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.

## Positive assoication between child allopaenting effort and number of older children

There was a statistically significant negative relationship between the number of older children in a household and the frequency of child alloparenting effort `r stats$malloparenting$OlderKids$str` when controlling for the sex of the child, the number of younger children, household income (log), the number of older children, the number of adults who provide childcare within the household, and the summed frequency of alloparenting from all other children in the household. As our model could not handle high values of the summed frequency of other child alloparenting, we removed `r sum(d2$OtherChildAlloparentingFreqN > 121, na.rm = TRUE)` outliers (`r sum(d2$OtherChildAlloparentingFreqN == 360, na.rm = TRUE)` of which had a value of 360 times per month and `r sum(d2$OtherChildAlloparentingFreqN == 420, na.rm = TRUE)` of which had a value of 420 timers per month) despite expecting these values to not be outliers in the broader Utilian community. See @tbl-alloparenting_dep_var for the corresponding regression table and @fig-effect_alloparenting_dep_var_plot for effects plots.

::: {#fig-effect_alloparenting_dep_var_plot}
```{r}
alloparenting_dep_var_plot
```

The effect of the number of younger children in the household on the frequency of a child's alloparenting effort. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-alloparenting_dep_var for the corresponding regression table.
:::
